---
title: Interaction Between the Effects of Sport Participation and BMI on Hard Drug
  Use
author: "Randall Boyes"
date: "January 31, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
#0.1. Required libraries

library(readr)
library(lme4)
library(dplyr)

#1.0. Read data file from CSV and convert to dataframe

HBSC_data <- read_csv("C:/Users/Student/Desktop/Data/HBSC Sport and Substance BMI/HBSC data.csv")

d <- as.data.frame(HBSC_data)

#2.0. Create 0/1 exposure variables as numeric

d$weed <- as.integer(d$weed)
d$sports <- as.integer(d$sports)
d$team <- as.integer(d$team)
d$ind <- d$sports - 2*d$team
d$chew <- as.integer(d$chew)
d$siblings <- as.integer(d$siblings)
d$region <- as.integer(d$`Region/Municipality`)
d$grade_cat <- if_else(d$Grade <= 9,9,10)

#3.0. Create Dummy variables for factor variables (there are missing data leaks in this code, needs fixing)

#3.1 BMI

d$bmifactor <- as.factor(d$`bmi classification`)
m <- dummy(d$bmifactor)
mdf <- as.data.frame(m)
mdf <- mutate(mdf, any_thin = xor(thin, `severely thin`))
d <- cbind(d,mdf)

#3.2 Family Structure

m2 <- as.data.frame(dummy(as.factor(d$family)))
colnames(m2)<-c("fam2","fam3","fam4")
d <- cbind(d,m2)

#3.3 Socioeconomic Status

m3 <- as.data.frame(dummy(as.factor(d$`Family well off`)))
colnames(m3)<-c("SESav","SESbad","SESfair","SESgood","SESgreat")
d <- cbind(d,m3)

#3.4 Ethnicity

d<-mutate(d, white <- ifelse(eth == "Caucasian",1,0))
d<-rename(d, "white" = "white <- ifelse(eth == \"Caucasian\", 1, 0)")

#3.5 Statscan Rurality

m4 <- as.data.frame(dummy(as.factor(d$`StatsCan Code`)))
colnames(m4) <- c("RURmed", "RURlarge","RURvsmall","RURsmall")
d <- cbind(d, m4)

#3.6 Rurality



#4.0 Drop unnecessary variables

d2 <- d[,c("Gender","team","ind","harddrug","siblings","obese","overweight","any_thin",
           "School","white","SESav","SESbad","SESfair","SESgood","SESgreat","fam2","fam3","fam4",
           "RURmed", "RURlarge","RURvsmall","RURsmall","grade_cat")]

#5.0 Stratify by gender

dg2 <- d2[d2$Gender == "Girl",]
db2 <- d2[d2$Gender == "Boy",]

#6.0 Adds a unique conseq. numeric ID variable for each school in the data set 

db2$SchoolID <- as.numeric(factor(db2$School, levels = unique(db2$School)))
dg2$SchoolID <- as.numeric(factor(dg2$School, levels = unique(dg2$School)))

#7.0 Missing data handling - currently complete case is implemented

db2 <- na.omit(db2)
dg2 <- na.omit(dg2)

```

## Model Specification

The working model for this paper is a sex-stratified Bayesian random intercepts logistic model. Hard drug use is modelled as a function of a binomial distribution with the logit of p determined by a linear combination of the exposures and confounding variables. Weakly regularizing normal priors are used for all parameters with the exception of the standard distribution of the school-level intercepts, which use a cauchy prior. Interaction effects are included for team sport participation, individual sport participation, and BMI.   

```{r model with priors}
m2.0 <- map2stan(
  alist(
    harddrug ~ dbinom(1,p),
    logit(p) <- a 
    + a_school[SchoolID] 
    + bTeam*team
    + bInd*ind
    + bSportInt*team*ind
    + bObese*obese
    + bTeamObeseInt*team*obese 
    + bIndObeseInt*ind*obese
    + bSportObeseInt*team*ind*obese
    + bOver*overweight
    + bTeamOverInt*team*overweight 
    + bIndOverInt*ind*overweight
    + bSportOverInt*team*ind*overweight
    + bSESgood*SESgood
    + bSESgreat*SESgreat
    + bSESfair*SESfair
    + bSESbad*SESbad
    + bfam2*fam2
    + bfam3*fam3
    + bfam4*fam4
    + bWhite*white
    + bSiblings*siblings
    + bRURmed*RURmed
    + bRURlarge*RURlarge
    + bRURvsmall*RURvsmall
    + bRURsmall*RURsmall
    + bGrade_cat*grade_cat,
    a_school[SchoolID] ~ dnorm(0,sigma_school),
    c(a,bTeam,bInd,bSportInt,bObese,bTeamObeseInt,bIndObeseInt,bSportObeseInt,bOver,bTeamOverInt,bIndOverInt,bSportOverInt) ~ dnorm(0,10),
    c(bSESgreat,bSESgood,bSESfair,bSESbad,bfam2,bfam3,bfam4,bWhite,bSiblings,bRURmed,bRURlarge,bRURvsmall,bRURsmall,bGrade_cat) ~ dnorm(0,10),
    sigma_school ~ dcauchy(0,1)
  ), 
  data = db2)
```

## Convergence of Markov Chains

Hamiltonian Markov Chain Monte Carlo using rStan was used to fit the model. The behavior of the chains is shown below.

```{r pressure, echo=FALSE}
plot(m2.0, pars = c("bTeam","bInd","bSiblings","bObese","bOver",
                    "bSESbad","bSESfair","bSESgood","bSESgreat","bfam2","bfam3","bfam4",
                    "bRURmed", "bRURlarge","bRURvsmall","bRURsmall","bGrade_cat","sigma_school"),max_rows = 6)
```

## Estimation of the effects

Estimation of the effects is shown below.

```{r}
plot(precis(m2.0))
```

```{posterior calculations, include=FALSE}
post <- extract.samples(m2.0)

#Combine these 1000 points to form predictions for all types of people 
#(not perfect - this code maintains the clusters from the sample rather than generalizing over average expected clusters)

post$pred_norm_ind <- post$bInd
post$pred_norm_team <- post$bTeam
post$pred_norm_both <- post$bInd + post$bTeam + post$bSportInt
post$pred_over_ind <- post$bInd + post$bIndOverInt + post$bOver
post$pred_over_team <- post$bTeam + post$bTeamOverInt + post$bOver
post$pred_over_both <- post$bInd + post$bTeam + post$bOver + post$bSportInt + post$bTeamOverInt  + post$bIndOverInt + post$bSportOverInt 
post$pred_obese_ind <- post$bInd + post$bIndObeseInt + post$bOver
post$pred_obese_team <- post$bTeam + post$bTeamObeseInt + post$bOver
post$pred_obese_both <- post$bInd + post$bTeam + + post$bObese + post$bSportInt + post$bTeamObeseInt  + post$bIndObeseInt + post$bSportObeseInt

```



## Plots

Plots of the estimation of effect are shown here:

```{r}
dens(post$pred_obese_ind, xlim = c(-1.5,1.5),col="#00FF00",ylim=c(0,4), xlab = "Beta",ask=FALSE)
dens(post$pred_obese_team, add=TRUE,col="#00FF00")
dens(post$pred_obese_both, add=TRUE,col="#00FF00")
dens(post$pred_over_ind, add=TRUE,col="#FF0000")
dens(post$pred_over_team, add=TRUE,col="#FF0000")
dens(post$pred_over_both, add=TRUE,col="#FF0000")
dens(post$pred_norm_ind, add=TRUE,col="#0000FF")
dens(post$pred_norm_team, add=TRUE,col="#0000FF")
dens(post$pred_norm_both, add=TRUE,col="#0000FF")
```

